---
title: "Week 12 IP"
author: "Ann Mberi"
date: "7/9/2021"
output:
  html_document: default
  word_document: default
---

## 1. INTRODUCTION

### 1.1 Defining the Question
My work is to identify which factors determine whether a user clicks on an ad or not.

### 1.2 Setting the Metric for Success
The project will be considered a success when I am able to identify what makes a user more likely to click on an ad.

### 1.3 Outlining the Context
A Kenyan entrepreneur has created an online cryptography course and would want to advertise it on her blog. She currently targets audiences originating from various countries. In the past, she ran ads to advertise a related course on the same blog and collected data in the process. She would now like to employ your services as a Data Science Consultant to help her identify which individuals are most likely to click on her ads.

### 1.4 Drafting the Experimental Design
1. Define the question, set the metric for success, outline the context, drafting the experimental design, and determining the appropriateness of the data.
2. Load the dataset and previewing it.
3. Check for missing and duplicated values and deal with them where necessary.
4. Check for outliers and other anomalies and deal with them where necessary.
5. Perform univariate and bivariate analysis.
6. Conclude and provide insights on how this project can be improved.



# Import Libraries 

install.packages("foreign")
library(foreign)
install.packages("car")
install.packages("Hmisc")
install.packages("reshape")
library(dplyr)

## Loading the Dataset

```{r}
df <- read.csv("http://bit.ly/IPAdvertisingData")
```


## Previewing the dataset 

```{r}
head(df) #The first 6 rows 
```


# checking the shape of the dataset
```{r}
dim(df)
```

## Tidying the Dataset

# Finding any null values in the data
```{r}
colSums(is.na(df))
```

# Lets check to see the specific number of missing values in our data

```{r}
sum(is.na(df))
```

We have no missing values 

# Checking for any duplicates 
```{r}
unique_items <- df[duplicated.default(df), ]
unique_items
```

There were no duplicated values in our data 


# checking for outliers precisely on the continuous variables

```{r}
cont_cols <- df[c('Daily.Time.Spent.on.Site','Age', 'Daily.Internet.Usage')]       
boxplot(cont_cols, main='BoxPlots')
```


# checking outliers for the Area Income 

```{r}
boxplot(df$Area.Income,xlab= 'Area.Income', main='boxplot')
```

There are outliers in the area income variable, this may be due to disparities in currencies between different countries


# PERFORMING EDA
## UNIVARIATE ANALYSIS

```{r}
summary(df)
library(psych)
describe(df)
```

# We calculate the means

```{r}
colMeans(df[sapply(df,is.numeric)])
```
 From our observation we can see 
 
 1. The time spent on site  on average is 65
 2. Average age of the user is 36 years
 3. Also we see that the average area income is 55000
 
 
# Mode 
# creating the function for mode
```{r}
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

apply(df,2,getmode)
```
# Obeservation 
1. The most common age of the users is 31 years
2. The most common gender in the data is male. 
3. The City with the most repeat users is Lisamouth
4. The country that's most repeated is Czech Republic


# getting the median of the various numeric variables
```{r}
apply(df[ , c(1:4, 7, 10)],2,median)

```

# Range 
```{r}
apply(df[ , c(1:4, 7, 10)],2,range)
```

# Quantile 
```{r}
apply(df[ , c(1:4, 7, 10)],2,IQR)
```

# VARIANCE 
```{r}
apply(df[ , c(1:4, 7, 10)],2,var)
```

# Standard Deviation 
```{r}
apply(df[ , c(1:4, 7, 10)],2,sd)
```

# Lets create a dataframe for those who clicked the ad 

```{r}
clicks <- df[df$Clicked.on.Ad==1,]
head(clicks)
```

# Bar plot
## We view per gender those who clicked ads

```{r}
gender_dist<- table(clicks$Male)
label<- c("female","male")
barplot(gender_dist,names.arg=label,main="gender distribution")
```

# We now view based on age distribution 
```{r}
age_dist<- table(clicks$Age)
barplot(age_dist,main="age distribution")
```

# Histogram 
# Lets view based on Area income 

```{r}
hist(clicks$Area.Income)
```

# Observations 
1. People aged 45 had the most clicks among all closely followed by those aged 36 and 38 years.
2. We had more females to  males counterparts in the number of ads clicked.
3. The areas that have an income between 40000 and 60000 have the most clicks on the ads

# Bivariate Analysis 
## Lets observe  correlation heatmap for the numerical columns in our dataset 

```{r}
cor(df[sapply(df,is.numeric)])
```

## Covariance

```{r}
cov(df[ , c(1:4, 10)])
```
As listed above there are varaibles where there are negative and positive covariances.The negative covariances indicate that as one variable increases,the second variable tends to decrease

## Correlation matrix

```{r}
library('corrplot')
```


```{r}
corrplot(cor(df[ , c(1:4, 10)]), method = 'number',type='lower',tl.srt = 0, main="correlation matrix" )
```
# Observations 

 * The 'daily internet usage' and 'daily time spent on site' have a negative strong correlation on the 'click on ad' variable.
* the area income also has a moderate negative correlation on the click on ad variable


# Scatter Plots 
### Daily time spent 
```{r}
plot(df$Clicked.on.Ad, df$Daily.Time.Spent.on.Site, xlab="Clicked on Ad", ylab="Daily time spent on site")
```

### Age

```{r}
plot(df$Clicked.on.Ad, df$Age, xlab="Clicked on Ad", ylab="Age")
```


### Daily time spent 

```{r}
plot(df$Daily.Time.Spent.on.Site, df$Age, xlab="Daily.Time.Spent.On.Site", ylab="Age")
```



### Area income 

```{r}
plot(df$Clicked.on.Ad, df$Area.Income, xlab="Clicked on Ad", ylab="Area Income")
```

# Conclusion
We built a profile of an individual most likely to click on the ad. That is a female between the age of 35-45 years. MOst probably from Australia, Ethiopia or Turkey.